---
title: "Homework#2"
author: "Delores Tang"
date: "10/14/2018"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(lattice)
library(MASS)
library(dplyr)
```


1. Imputing age and gender
```{r, include = F}
#Import datasets
BestInc <- read.csv("C:/Users/delor/persp-analysis_A18/Assignments/A2/BestIncome.txt", header=FALSE)
SurvInc <- read.csv("C:/Users/delor/persp-analysis_A18/Assignments/A2/SurvIncome.txt", header=FALSE)
IncIntel <- read.csv("C:/Users/delor/persp-analysis_A18/Assignments/A2/IncomeIntel.txt", header=FALSE)

# Rename variables
names(BestInc) <- c("L_inc","C_inc","Height","Weight")
names(SurvInc) <- c("T_inc","Weight", "Age", "Gender")
names(IncIntel) <- c("year","gre","salary")
```

(a) In order to impute age and gender from SurvIncome to BestIncome, I would like to fit a generalized linear model to variables in SurvIncome.
```{r, tidy = TRUE}
#Linear Regression in SurvIncome
Agelm <- lm(Age ~ Weight + T_inc, data = SurvInc)
summary(Agelm)
Genderlm <- lm(Gender ~ Weight + T_inc, data = SurvInc)
summary(Genderlm)
```


Therefore, the equations that could be used to impute Age and Gender to BestIncome will be: $$\text{Age} = 44.21 - 6.722\times 10^{-3} \times \text{Weight} + 2.52 \times 10^{-5} \times\text{Total Income}$$ and $$\text{Gender} = 3.761 - 1.953\times 10^{-2} \times \text{Weight} -5.25 \times 10^{-6} \times\text{Total Income.}$$  


(b) Using the equations obtained from question (a), I imputed Age and Gender to BestIncome based on the SurveyIncome dataset. Since gender has to be binay, I set all imputed gende that is greater than 0.5 as 1, and others as 0. Also, I assumed that the sum of labor and capital income in BestIncome data is equivalent to the variable Total Income in SurveyIncome data.

```{r, tidy = TRUE}
BestInc$Age <- 4.421e+01 -6.722e-03 * BestInc$Weight + 2.520e-05 * (BestInc$L_inc + BestInc$C_inc)
BestInc$Gender <- 3.761e+00 -1.953e-02 * BestInc$Weight -5.250e-06 * (BestInc$L_inc + BestInc$C_inc)

# Since Gender has to be binary
for (i in 1:length(BestInc$Gender)){
  if (BestInc$Gender[i] < 0.5){
       BestInc$Gender[i] <- 0
  }else{
       BestInc$Gender[i] <- 1
  }
}
```
  
(c) For age,  
mean: 44.89,    sd = 0.219,     min = 43.98,    max = 45.70,    no. of observation = 10000;  
For Gender,  
mean: 0.4614    sd = 0.499,     min = 0,        max = 1,        no. of observation = 10000.
```{r, tidy = TRUE}
summary(BestInc$Age)
summary(BestInc$Gender)
sd(BestInc$Age)
sd(BestInc$Gender)
length(BestInc$Age)
length(BestInc$Gender)
```
  
(d) Correlation matrix:
```{r, tidy = TRUE}
res <- cor(BestInc)
round(res, 2)
```
  
  
2.(a) Coefficients for intercept: 89541.293,    Std.Error: 878.764  
      Coefficients for GRE score: -25.763,      Std.Error: 1.36
```{r, tidy = TRUE}
salmod1 <- lm(salary ~ gre, data = IncIntel)
summary(salmod1)
```
  
(b) Due to the change in GRE's scoring system, a system drift on the data occurred at the year 2010-2011. Therefore, as indicated by the scatterplot, people's GRE scores dropped significantly due to this change. 

```{r, tidy = TRUE}
#Scatter plot of Gre scores vs. Graduation year
xyplot(gre ~ year, xlab = "Graduation year", ylab = "Gre Quantitative Scores", data = IncIntel)
```

To accurately test our hypothesis, we would have to rescale GRE scores to eliminate the effect of the data drift by using the percentile a person gets in the year he takes GRE test, instead of the general GRE quantitative score, to indicate his or her academic performance on a GRE test. 
```{r, tidy = TRUE}
#Define a function that extract a list for each year's gre scores
grefunc <- function(yr) {
  grelist <- list()
  for (i in (1:nrow(IncIntel))) {
    if (IncIntel$year[i] == yr) {
      grelist <- append(grelist, IncIntel$gre[i])
    }
  }
  return(grelist)
}

#Make a dictionary that sort each year's participants' gre scores
gredic <- list()
for (yr in (2001:2013)){
  grelist <- grefunc(yr)
  gredic <- append(gredic, list(grelist))
  
}
```


```{r, tidy=TRUE}
#Create a new variable gre_perc that indicates the participant's percentile GRE in that year
for (yr in (2001:2013)) {

  for (i in (1:nrow(IncIntel))) {
    if (IncIntel$year[i] == yr) {
      dicyr <- IncIntel$year[i] - 2000 
      gredic_yr <- unlist(gredic[dicyr]) 
      gredic_yr_rank <- ecdf(gredic_yr) 
      IncIntel$gre_perc[i] <- gredic_yr_rank(IncIntel$gre[i])
    }
  }  
}

xyplot(IncIntel$gre_perc ~ IncIntel$year, xlab = "Graduation year", 
       ylab = "Percentile GRE Quantitative Scores")
```


(c) 
```{r, tidy = TRUE}
#Scatterplot of the original salary vs. year
xyplot(salary ~ year,  xlab = "Graduation year", ylab = "Salary after 4 years of graduation", data = IncIntel)
```
  
  As indicated by the scatter plot, salary inflates gradually over the years. In that case, years is a stationarity that could confound our hypothesis testing. Therefore, we could calculate the average annual growth rate of income, and balance out this annual rate of salary growth by dividing a person's salary by the growth rate to the power of time (in years).
    
```{r, tidy = TRUE}
#Control for stationarity
##Average income per year
avg_inc <- aggregate(select(IncIntel,salary), list(IncIntel$year), mean)
avg_inc

##Average growth rate
avg_growth <- mean((avg_inc$salary[13:2] - avg_inc$salary[12:1]) / avg_inc$salary[12:1])
IncIntel$ctrSalary <- IncIntel$salary/((1 + avg_growth)**(IncIntel$year - 2001))
xyplot(IncIntel$ctrSalary ~ IncIntel$year)
```

(d) Estimated coefficients:  
Intercept: 61643.2,     Std.Error: 455.9;  
GRE percentiles: -411,  Std.Error: 782.0.  
  
Now the coefficient for GRE percentile, unlike that in (a), turned to be insignificant (p > 0.05). 
  
Due to the change we made in (b), the estimated values of both coefficients varied signifiantly as we adjusted GRE variable to a new GRE percentile variable that ranges from 0 to 1. 
  
Similarly, after controlling for system drift in GRE scores and the stationarity of time, the results of linear regression model suggests that GRE quantitative score is not a significant predicting factor of one's salary after 4 years of graduation. It suggests that the alternative hypothesis is likely to be rejected. 

```{r, tidy = TRUE}
adjustLm <- lm(ctrSalary ~ gre_perc, data = IncIntel)
summary(adjustLm)
```